---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# sidrar

[![CRAN status](https://www.r-pkg.org/badges/version/sidrar)](https://CRAN.R-project.org/package=sidrar)
[![R-CMD-check](https://github.com/rpradosiqueira/sidrar/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rpradosiqueira/sidrar/actions/workflows/R-CMD-check.yaml)

`sidrar` provides direct access from R to aggregate data and metadata
published by the Brazilian Institute of Geography and Statistics (IBGE).
SIDRA stands for *Sistema IBGE de Recuperação Automática*.

## Installation

Install the released version from CRAN:

```{r, eval = FALSE}
install.packages("sidrar")
```

Install the development version from GitHub with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("rpradosiqueira/sidrar")
```

## Main functions

The original three entry points remain unchanged:

- `search_sidra()` searches the current official aggregate catalog.
- `info_sidra()` lists the parameters available for a table.
- `get_sidra()` retrieves the selected observations.

Table codes are returned as the names of the `search_sidra()` result:

```{r, eval = FALSE}
search_sidra("IPCA")
info_sidra(7060)
```

An additive workflow is available for programmatic discovery, planning, and
larger requests:

- `sidra_catalog()`, `sidra_metadata()`, `sidra_periods()`, and
  `sidra_locations()` return structured discovery data.
- `sidra_query()` constructs a request without downloading values, and
  `sidra_plan()` reports the cardinalities that can be determined offline.
- `sidra_split()` creates explicit disjoint batches and `sidra_collect()`
  retrieves them sequentially with schema validation.

## Discover and plan

The structured discovery functions keep identifiers as character strings and
return stable base R objects:

```{r, eval = FALSE}
catalog <- sidra_catalog()
metadata <- sidra_metadata(7060)
periods <- sidra_periods(7060)
brazil <- sidra_locations(7060, "N1")
```

Build and inspect a query before downloading values:

```{r, eval = FALSE}
query <- sidra_query(
  x = 7060,
  variable = 63,
  period = sprintf("2024%02d", 1:12),
  geo = "City",
  geo.filter = list(City = 5002704),
  classific = "c315",
  category = list(7169)
)

query$url
sidra_plan(query)
```

## Retrieve data

This example requests the monthly IPCA for the general index in Campo
Grande, Mato Grosso do Sul, over the 12 most recent periods:

```{r, eval = FALSE}
library(sidrar)

ipca <- get_sidra(
  x = 7060,
  variable = 63,
  period = c(last = 12),
  geo = "City",
  geo.filter = list(City = 5002704),
  classific = "c315",
  category = list(7169)
)
```

You may also pass either a relative API path or a complete official HTTPS
URL. A request containing `/h/n` is returned without consuming its first
observation as a header:

```{r, eval = FALSE}
ipca_brazil <- get_sidra(
  api = paste0(
    "https://apisidra.ibge.gov.br/values/",
    "t/7060/n1/all/v/63/p/last/c315/7169"
  )
)
```

## Preserve special values

By default, `Valor` is numeric for compatibility with earlier releases.
SIDRA also uses symbols such as `"-"`, `"X"`, `".."`, and `"..."`.
Use `value_type = "character"` to preserve them in `Valor`, or
`value_type = "both"` to append `Valor_raw` while retaining numeric
`Valor`:

```{r, eval = FALSE}
data <- get_sidra(
  api = "/t/1849/n3/all/v/811/p/2018/c12762/all",
  value_type = "both"
)
```

## Network behavior and cache

The legacy functions continue to query current IBGE services on every call.
The structured discovery functions have an opt-in disk cache; value responses
are never cached automatically:

```{r, eval = FALSE}
metadata <- sidra_metadata(7060, cache = TRUE)
sidra_cache_info()
sidra_cache_clear()
```

Cache entries expire after 30 minutes by default. Use `refresh = TRUE` to
bypass and replace an entry. Save collected values explicitly when a
reproducible snapshot is required. Requests use a timeout and limited retries
for transient failures; customize them with:

```{r, eval = FALSE}
options(
  sidrar.timeout = 120,
  sidrar.retries = 4
)
```

Regular package tests are offline. Live API smoke tests run separately on a
small set of queries to detect availability and schema changes.

### Requests above the SIDRA value limit

When SIDRA rejects a request because it exceeds the service's per-request
value limit, `get_sidra()` raises a `sidrar_limit_error`. Its message reports
the requested count, the current limit, and the minimum number of calls. The
condition also inherits from `sidrar_http_error`, so existing error handlers
continue to work.

`sidra_split()` partitions one explicit dimension without splitting category
sums or changing the geographic level. A geographic filter is splittable only
when the query requests one non-Brazil level; otherwise unchanged territorial
levels could overlap between calls. Relative periods, ranges, and embedded
comma lists must first be expanded into one explicit code per vector element.
`sidra_collect()` runs the resulting queries sequentially and refuses to
combine incompatible schemas:

```{r, eval = FALSE}
query <- sidra_query(
  x = table_code,
  variable = variable_code,
  period = period_code,
  geo = "City",
  classific = classification_code,
  category = list(category_codes)
)

batches <- sidra_split(query, by = "category", size = 8)
data <- sidra_collect(batches, provenance = TRUE)
sidra_provenance(data)
```

The appropriate group size depends on the cardinality of every selected
dimension. Automatic implicit splitting remains disabled.

### Geographic identifiers

Keep geographic identifiers as character strings. SIDRA neighborhood codes
(`geo = "Neighborhood"`) identify its neighborhood territorial level; they
are not census tract identifiers and must not be joined directly to census
tract geometries without an official correspondence.

Structured queries also support official territorial views and extinct
territorial units:

```{r, eval = FALSE}
sidra_query(1612, geo_view = 44, classific = character())
sidra_query(
  1612,
  geo = "State",
  geo.filter = list(c(20, 34)),
  include_extinct = TRUE,
  classific = character()
)
```

For more examples, see the
["Introduction to sidrar"](https://CRAN.R-project.org/package=sidrar/vignettes/Introduction_to_sidrar.html)
vignette and the
[official SIDRA API documentation](https://apisidra.ibge.gov.br/home/ajuda).
